\documentclass[11pt,a4paper]{article}
\usepackage{amsmath, amssymb, fullpage, mathrsfs, bm, pgf, tikz}

\begin{document}

\title{Binary quadratic form + $n$-th power residues.}
\author{Anzo Teh}
\date{28 July 2016}
\maketitle

\begin {enumerate}
\item \emph {JOM 2013, N5.} Let $p$ be an odd prime such that $2^p+1|p^p+1$. Prove that any prime divisor of $2^p+1$ other than $3$ is greater than $6p$.\\
\emph {Generalization.} For any prime number $p$ we have: each prime divisor of $\gcd(2^p+1,p^p+1)$ is equal to $3$ or is greater than $6p$.

\item\emph {APMO 2014, \# 3.} Find all positive integers $n$ such that for any integer $k$ there exists an integer $a$ for which $a^3+a-k$ is divisible by $n$.

\item\emph{APMO 2012, \# 3.}  Determine all the pairs $ (p , n )$ of a prime number $ p$ and a positive integer $ n$ for which $ \frac{ n^p + 1 }{p^n + 1} $ is an integer.

\item\emph{IMO 2012, N6.} Let $x$ and $y$ be positive integers. If ${x^{2^n}}-1$ is divisible by $2^ny+1$ for every positive integer $n$, prove that $x=1$.

\item\emph {IMO 2012, N8.} Prove that for every prime $p>100$ and every integer $r$, there exist two integers $a$ and $b$ such that $p$ divides $a^2+b^5-r$.

\item\emph {IMO 2011, N8.} Let $k \in \mathbb{Z}^+$ and set $n=2^k+1.$ Prove that $n$ is a prime number if and only if the following holds: there is a permutation $a_{1},\ldots,a_{n-1}$ of the numbers $1,2, \ldots, n-1$ and a sequence of integers $g_{1},\ldots,g_{n-1},$ such that $n$ divides $g^{a_i}_i - a_{i+1}$ for every $i \in \{1,2,\ldots,n-1\},$ where we set $a_n = a_1.$

\item\emph {IMO 2010, N4.} Let $a, b$ be integers, and let $P(x) = ax^3+bx.$ For any positive integer $n$ we say that the pair $(a,b)$ is $n$-good if $n | P(m)-P(k)$ implies $n | m - k$ for all integers $m, k.$ We say that $(a,b)$ is $very \ good$ if $(a,b)$ is $n$-good for infinitely many positive integers $n.$

    (a) Find a pair $(a,b)$ which is 51-good, but not very good.\\
    (b) Show that all 2010-good pairs are very good.

\item\emph {IMO 2008, \# 3.} Prove that there are infinitely many positive integers $ n$ such that $ n^{2} + 1$ has a prime divisor greater than $ 2n + \sqrt {2n}$.


\end{enumerate}

\end{document}